### 2012-07-10 paper.tex

\documentclass[letterpaper, 12pt]{article}
%%\documentclass[letterpaper, 11pt, twocolumn]{article}
\usepackage{url}
\usepackage{ctable}
\usepackage{overcite}
\usepackage{amsmath,amssymb}      % for \AA etc
\usepackage[onehalfspacing]{setspace}
\usepackage{graphicx}
\usepackage{fullpage, times}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{longtable}

\DeclareGraphicsExtensions{.pdf,.png}
\onehalfspacing

\bibliographystyle{jcics}

\usepackage{anysize}
%\marginsize{1in}{2in}{1in}{1in}
%\marginparwidth 1.75in
%\newcommand{\mnote}[1]{\marginpar{\raggedright{\footnotesize\textbf{#1}}}}

\def\imagetop#1{\vtop{\null\hbox{#1}}}

%%%
%%% Don't change the size of the figures as they are defined for
%%% a single column environment, which is what we will be submitting
%%% as
%%%

\begin{document}
\title{Exploring Uncharted Territories - Predicting Activty Cliffs in
  Structure-Activity Landscapes}
\author{Rajarshi Guha${}^{\ddagger}$\\
NIH Center for Advancing Translational Sciences \\ 9800 Medical Center Drive  Rockville, MD 20850}
\date{}

\maketitle
\begin{abstract}
  The notion of activity cliffs is an intuitive approach to characterizing structural features that
  play a key role in modulating biological activity of a molecule. A variety of methods have been
  described to quantitatively characterize activity cliffs, such as SALI and SARI. However, these
  methods are primarily retrospective in nature; highlighting cliffs that are already present in the
  dataset. The current study focuses on employing a pairwise characterization of a dataset to train
  a model to predict whether a new molecule will exhibit an activity cliff with one or more members
  of the dataset. The approach is based on predicting a value for pairs of objects rather than the
  individual objects themselves (and thus allows for robust models even for small SAR datasets). We
  extracted structure-activity data for several ChEMBL assays and developed random forest models to
  predict SALI values, from pairwise combinations of molecular descriptors. The models exhibited
  reasonable RMSE's though, surprisingly, performance on the more significant cliffs tended to be
  better than on the lesser ones. Our results indicate that the predictive models are able to
  prioritize molecules in terms of their ability to activity cliffs and can be used as a way to
  extend an observed activity landscape to identify molecules that could lead to significant
  improvements in activity. Conversely, the method can be used to deprioritize molecules that are
  predicted to show an activity cliff with a potent molecule already in the dataset.
\end{abstract}

\section{Introduction}

The landscape paradigm for SAR data was first proposed 20 years ago \cite{Johnson:1990ys} and has
recently seen a resurgence with a number of studies describing new ways to quantify and visualize
activity landscapes. When SAR data is viewed as a landscape, with the X-Y plane representing
structural characteristics (which will usually be a 2-dimensional representation of a
multi-dimensional descriptor space) and the Z-axis representing the observed activities, one can
identify two broad types of regions on the landscape - smooth rolling regions, corresponding to set
of molecules exhibiting continuous SAR (i.e., similar structures and similar activities) and rough,
gorge-like regions (i.e., very similar structures, but large differences in activity) corresponding
to molecules that exhibit SAR discontinuity. The latter have also been term activity cliffs
\cite{Maggiora:2006aa}. From a medicinal chemistry point of view, the latter regions of a landscape
can be the most interesting as they can provide insight into structural features that are key to
improving (or conversely reducing) potency. There is a rich history of methods that have correlated
structural differences with corresponding differences in activity -- matched molecular pairs
\cite{Leach:2006aa}, SAS maps \cite{Maggiora:2011kx} and more recently SALI \cite{Guha:2008aa} and
SARI \cite{Peltason:2007aa}. Both SALI and SARI focus on numerically characterizing a structure
activity landscape. The former is defined for a pair of molecules as
\begin{equation}
  \label{eq:1}
  S_{i,j} = \frac{|A_i - A_j|}{1 - sim(i,j)}
\end{equation}
where $A_i$ and $A_j$ represent the observed activities of molecules $i$ and $j$, and $sim(i,j)$
represents the structural similarity between the two molecules (usually based on some form of
fingerprint similarity). Using the SALI metric, one can take a collection of $n$ molecules and
represent them as an $n \times n$ matrix of SALI values - larger values representing more
significant activity cliffs. The SARI approach is based on a score defined as
\begin{equation}
  \label{eq:2}
  \mathrm{SARI} = \frac{1}{2} \left(\mathrm{score}_{cont} + (1 -
    \mathrm{score}_{disc}) \right)
\end{equation}
where the individual score terms are derived on the basis of potency and pairwise similarities. The
reader is referred to Ref. \citen{Peltason:2007aa} for a detailed discussion of this approach.

In either case, one can take the numerical values and visualize them in a variety of ways ranging
from heatmaps of SALI matrices to network representations \cite{Wawer:2008aa,Guha:2008aa}. These
visualizations then allow the user to explore the landscape, quickly identifying a range of activity
cliffs, which an then be examined in detail. Apart from identifying individual activity cliffs, a
variety of other SAR constructs, such as ``activity ridges'' \cite{Vogt:2011bs} and multitarget
landscapes\cite{Dimova:2011fk} can also be identified and characterized.

A feature common to all, recently published work on activity landscapes is that they are primarily
retrospective. That is, the methodologies developed are used to analyze SAR datasets for which
activities have already been experimentally obtained. For example, using the SALI, one can
characterize SAR patterns in a dataset but does not provide insight into whether a new molecule may
be part of an activity cliff, with respect to the original dataset. A number of applications have
attempted to extract SAR rules based on the landscape (e.g., similarity potency trees
\cite{Wawer:2010fv} and multi-target landscape analyses \cite{Hu:2010kl}) or directly identify
structural modifications that lead to activity cliffs \cite{Wassermann:2010oq, Peltason:2009vn}.

\subsection{Motivation}
\label{sec:motivation}
The preceding discussion highlights the utility of retrospective analyses of SAR data using the
activity landscape paradigm. But equally, if not more, interesting is determining whether a new,
untested molecule might be an activity cliff in the context of the original dataset. For the
remainder of this work we focus on the use of SALI to quantify activity landscapes. More
specifically, the ability to \emph{predict} SALI values would be useful as it would allow us to both
fill in empty regions of an activity landscape as well as extend a structure-activity
landscape. Note that this approach to expanding the extent of a SAR dataset does not lend itself to
scaffold hopping since the premise of scaffold hopping is that one generates new cores, which differ
substantially from the starting structure.

In traditional QSAR modeling approaches, one simply predicts the activity of a new molecule and
would then evaluate the SALI (or SARI or some other measure) to determine whether the molecule leads
to an activity cliff. However, the fact that an activity cliff represents a SAR
discontinuity\cite{Maggiora:2006aa} implies that most statistical and machine learning methods will
be unlikely to predict very different activities for two structurally similar molecules. In other
words, a new molecule, similar to a subset of the training set, will tend to have a predicted value
that is similar to those molecules, rather than a drastically different value.

An alternative approach, that is the focus of this paper, is to directly predict SALI values for
pairs of molecules. Thus rather than predict individual activities, we predict SALI values for pairs
of molecules. This approach is somewhat similar to the SPREAD method \cite{Scheiber:2009fk} which
identified substructures that were predictive of activity differences. However, our solution
considers both activity differences and structural similarities. As a result, instead of ranking
compounds in terms of their predicted activity, we instead rank a compound in terms of its predicted
SALI; i.e., its predicted ability to exhibit an activity cliff when paired with other molecules in
the dataset. This approach could be useful when deciding how far to extend an analog series as well
as prioritizing scaffolds for further study.

This does not completely alleviate the problem of discontinuities, since SALI values are infinite
when the $T_c$ is 1.0.  However, predicting SALI values allows us to work with smaller datasets
(since the objects to predict are pairs of molecules), that would ordinarily lead to unreliable
models if we were working with activities. Of course, this means that the approach is not practical
for very large datasets.

The paper is organized as follows. Section \ref{sec:datasets} describes the datasets used in this
study. Section \ref{sec:methodology} presents the methodology we employ to predict activity cliffs
and Section \ref{sec:applications} discusses the results of the predictive models. Finally, Sections
\ref{sec:discussion} and \ref{sec:conclusions} discusses some of the issues underlying this approach
and possible extensions of this work.

\section{Datasets}            
\label{sec:datasets}
For this study we considered a number of datasets, which are summarized in Table \ref{tab:data}. The
Cavalli dataset was employed in our previous studies consisted of 30 molecules studied by Cavalli et
al\cite{Cavalli:2002aa} as possible hErg inhibitors using a pharmacophore modeling approach. The
remaining datasets were obtained from ChEMBL. All three assays involved direct binding to a human
target and we considered the subset of molecules in each assay that had non-censored experimental
values. The Costanzo dataset\cite{Costanzo:2005uq} consisted of 60 $\alpha$-ketoheterocyclic
inhibitors of $\alpha$-thrombin. The reported $\mathrm{IC_{50}}$ values ranged from 3 nM to 82
$\mu$M. The Kalla dataset\cite{Kalla:2006kx} consisted of 38 8-(C-4-pyrazolyl) xanthines, identified
as antagonists of the A2B adenosine receptor. The reported $K_i$ values ranged from 0.9 nM to 42
$\mu$M. Finally, the Dai dataset\cite{Dai:2007kl} consisted of 44 3-aminoindazole derivatives
studied for their ability to inhibit the VEGF and PDGF receptor families, with $\mathrm{IC_{50}}$
values ranging from 3 nM to 12~$\mu$M.

% To obtain the remaining datasets we mined the
% ChEMBL database to identify assays with the following constraints:
% cntaining between 30 and 500 molecules, direct binding to a human
% protein target, a data type of $K_i$ or $\mathrm{IC_{50}}$ and a
% reported confidence score of 8 or more. This identified 780
% assays. For each assay we considered those molecules with non-censored
% measured values and identified those assays for which at least one
% pair of molecules exhibited 1000-fold (or better) difference in
% activity. This led to a final set of 347 assays. From this set we
% considered three assays for three different targets

\section{Methodology}
\label{sec:methodology}
As noted above, the problem of identifying activity cliffs involving new molecules can be reduced to
predicting the SALI value for the new molecule and a pre-existing molecule. Note that this does not
result in an activity prediction for the new molecule; rather, it allows us to rank a set of new
molecules in terms of their predicted ability to exhibit a significant activity difference from one
or more of the molecules in the training set.

Given a training set of $N$ molecules, we generate a new training set
of $\frac{N(N-1)}{2}$ objects, where each object is a pair of
molecules from the original training set. The dependent variable for
each pair $i,j$, is the SALI value, $S_{i,j}$. SALI values were
evaluated using the 1051-bit BCI keyed fingerprints or the CDK
1024-bit path fingerprints. Based on the
definition of SALI, it is possible that a pair of molecules have a
$T_c = 1.0$ resulting in infinite values. For such cases, we replaced
the infinite value with the highest non-infinite SALI value for that
dataset.

The next step is to generate a set of independent variables. Since the
new dataset consists of pairs of molecules, we consider the
descriptors for the resultant objects as a function of the descriptor
values of the individual molecules. For an object, representing the
$i$'th and $j$'th molecules, its descriptor vector can be taken as the
arithmetic mean of the descriptor vectors of the individual
molecules. We denote this aggregation function as
$f_{\textrm{mean}}$. Alternative functions that were investigated included
the absolute difference of the individual descriptors, denoted by
$f_{\textrm{diff}}$ and the geometric mean of the individual
descriptor vectors, denoted by $f_{\textrm{geom}}$.

Given the independent and dependent variables for the pairwise
dataset, we can now proceed to model development. For this study we
focused on the use of random forest models\cite{Breiman:1984aa}. This
was motivated by the fact that such models can be proven to not
overfit and the fact that the algorithm implicitly performs feature
selection. As a result, this allows us to forgo an explicit feature
selection step and work directly with the descriptor pool (after
removal of correlated and constant descriptors). Furthermore, there is
no reason, \emph{a priori} to assume that the underlying SARs are
linear. A random forest model, being an algorithmic
approach\cite{Breiman:2001nx} (as opposed to a distributional one such
as linear regression) makes no such assumptions. We employed the  
implementation of random forest from the \texttt{randomForest}
package, in R 2.11.0\cite{r}, using the defult settings.

In this work, we employed the CDK to evaluate 109 2D and
constitutional descriptors for each individual molecule. For each
dataset, we performed objective feature selection by removing
descriptors with constant or near-constant values followed by removal
of descriptors that are highly correlated with others (using an $R^2$
cutoff of 0.8). The size of the final descriptor pools is summarized
in Table \ref{tab:data}.

One feature of this approach is that one can build relatively robust
models even for datasets of small size - say, 20 molecules. Of course,
a large training set is just one aspect of a reliable model and other
considerations such as diversity, descriptor selection and so on still
play an important role.

\subsection{Is structural information duplicated?}
\label{sec:are-we-duplicating}

One question that arises from the preceding discussion is the
possibility that we are duplicating information between the dependent
and independent variables. More specifically, both the dependent
variable (i.e., the SALI values) and the aggregated descriptors
characterize the molecules' structure. One might therefore ask whether
models built using such data perform over-optimistically. Given the
nature of the descriptors used in the indepdent and dependent
variables (the former being based on multiple atom and bond features
and the latter derived from purel topological paths), we believe that such a
phenomenon is unlikely.  Fig.~\ref{fig:desccor1} displays a histogram
of the pairwise Pearson correlations between the SALI values and each
of the descriptor values for the Cavalli\cite{Cavalli:2002aa} dataset,
using the different aggregation functions described above. It is
evident, that the highest $R^2$, between any of the descriptors and
the SALI values is less than 0.15. Similar behavior was observed for
all the datasets used in this study. These observations
suggest that the problem of including correlated structural
information simultaneously in the dependent and independent variables
is minimal.

\section{Results}
\label{sec:applications}
We first consider the application of the activity cliff prediction
methodology to the Cavalli dataset. We developed three random forest
models, using the three aggregation descriptor aggregation functions
described in Section
\ref{sec:methodology}. Fig.~\ref{fig:rfmodel-cavalli} presents the
predicted versus observed SALI values from the three models and Table
\ref{tab:rfmodel-results} summarizes the performance metrics for the
three models. Overall the different aggregation functions do not
differ dramatically in terms of the final model performance, though
$f_{\textrm{mean}}$ appears to lead to the best performance.

While the models are certainly not the most accurate models, the
RMSE's and $R^2$ values are reasonable. Note that the predictions on
the low end of the SALI values are not as important as those at the
high end - simply because low SALI values correspond to small activity
cliffs, which are likely not very interesting. Given that observation,
it is encouraging to observe that all three models perform relatively
well at the higher end of the SALI spectrum. For this example, the
three most significant activity cliffs are in fact not very
significant activity cliffs in an absolute sense - the $T_c$ for the
the three pairs are just 0.2, 0.30 and 0.29, though the activity
differences were 4.93, 5.0 and 5.11 log units respectively.

We observed that the use of the geometric mean did not lead to models
that were any different from those generated using the arithmetic mean
as the aggregation function. This was also true of the other datasets
and so the following discussion ommits the results obtained when using
$f_{\textrm{geom}}$. Fig.~\ref{fig:rfmodel-chembl} summarizes the
models built on the ChEMBL datasets using $f_{\textrm{diff}}$ and
$f_{\textrm{mean}}$ as the aggregation functions. Table
\ref{tab:rfmodel-results} reports the model statistics.  While the
differences in model performance do not vary significantly with the
aggregation function used, we see that for the Costanzo and Kalla
datasets, $f_{\textrm{diff}}$ leads to a slightly better model than
when using $f_{\textrm{mean}}$, though for the Dai dataset, the
difference is negligable.

Given the relatively small differences between models based on
different aggregation functions, we arbitrarily chose
$f_{\textrm{mean}}$ as the aggregagtion functin for subsequent
analyses. Figs.~\ref{fig:costanzo-pred}, \ref{fig:kalla-pred} and
\ref{fig:dai-pred} summarize the performance of the models built using
the reported activity data and the pairwise SALI values for the
Costanzo, Kalla and Froimowitz datasets, respectively. As with the
Cavalli dataset, the SALI models exhibit more variance for small SALI
values (i.e., less significant cliffs) but appear to perform better at
the the higher end of the SALI values. However in all three cases, the
most significant cliffs exhibit significant variance. In comparison,
the models developed on the reported activity values are relatively
poor. This can be partially ascribed to the small sizes of the
dataset.

\subsection{Extending a landscape}
\label{sec:extending-landscape}

Clearly, modeling the pairwise SALI values leads to reasonable models
the datasets discussed here. However, the primary goal of this
approach is to extend the the structure-activity landscape for a set
of molecules. In other words, can one use the model to identify
compounds that will exhibit an activity cliff with one or more
molecules in the dataset. Given a new molecule, one can evaluate the
descriptors for this molecule and then evaluate the pairwise
descriptor values with each member of the training set. Thus, given a
new molecule, we must obtain predictions for $n$ ``pairwise objects'',
where $n$ is the number of molecules in the training set. Given these
$n$ predictions one can proceed to identify training set molecules
with which the new molecule might exhibit a cliff.

We first performed this analysis with the Kalla dataset. Since
predictions of new molecules involve $n$ predictions per molecule, we
removed three molecules from the dataset and rebuilt the random
forest model. The hold out molecules, were specifically selected since
they displayed significant activity cliffs with various members of the
training set. We then evaluted the pairwise descriptor values for
these three with the training set molecules and obtained predictions
of the log(SALI) values. Fig.~\ref{fig:tset-pset-kalla}A
summarizes the performance of the model on the training set as well as
for the prediction set. From a numerical point of view, the
performance of the model on he prediction set degrades somewhat (RMSE
of 0.51 versus 0.41 for the training set). As noted above, the
variance of the prediction is higher at the lower scale of log(SALI)
values. However, at the higher end of the scale there a number of
pairs that are well predicted. Yet, it is clear that a number of
relatively significant activity cliffs have been underestimated.

Given that large SALI values can arise due to high structural
similarity, even when the activity difference is small, simply
examining log(SALI) values is not comletely informative. Specifically,
we are interested in the predictions for the actual activity cliffs,
where both the difference in activity and the structural similarity is
very high. Fig.~\ref{fig:tset-pset-kalla}B visualizes this
information. Each point coresponds to a pair of molecules, which are
shaded by the absolute prediction residual for that pair. Thus, the
``true'' cliffs are represented by points lying towards the bottom
right corner of the plot. In this case we see that there are a number
of pairs of molecules with $1 - T_C \leq 0.1$ and 100-fold or better
difference in activity. While many of these pairs are associated with
a high prediction residual, there are a number with low to medium
residuals.

Fig.~\ref{fig:kalla-comparison} displays the structures of the three
hold out molecules (top row) and a selection of structures, which
which these hold out molecules were predicted to exhibit an activity
cliff. For each molecule, the $K_i$ value is listed along with the
absoute residual value (log(SALI) units). For molecules
\textbf{349288} and \textbf{349699}, the predicted cliffs are
relatively accurate, as evidenced by the low residuals. For
\textbf{349138} on the other hand, most of the residuals were were
relatively high. While we have highlighted a number of cliffs,
Fig.~\ref{fig:tset-pset-kalla}C makes it clear that the predictions
for this molecule were, on average, worse than for the other two
molecules.

Fig.~\ref{fig:tset-pset-dai} summarizes the results for the same set
of analyzes described above, applied to the Dai dataset. As before we
removed three molecules (top row of Fig.~\ref{fig:dai-comparison}) and
rebuilt the random forest on the paired cases with the remaining 41
molecules. We then predicted the pairwise SALI values for the three
hold-out molecules with the remaining 41 molecules (i.e., the
prediction set). A plot of the predicted versus observed log(SALI)
values for the training and prediction sets are displayed in
Fig.~\ref{fig:tset-pset-dai}A. As before the model performs poorly on
the lesser cliffs, but improves for the more significant cliffs (RMSE
for the training set and prediction sets were 0.37 and 0.43 log units
respectively). However, the 2 most significant cliffs are
under-estimated. However, in comparison to the Kalla dataset, the
distribution of absolute residuals aggregated by the hold-out molecule
(Fig.~\ref{fig:tset-pset-dai}C) are reatively similar, with a median
absolute residual of less than 0.3 log units.  For this dataset, the
number of ``real'' cliffs is relatively low as shown in
Fig.~\ref{fig:tset-pset-dai}B. Interestingly, the compound pairs that
display moderate to significant cliffs are relatively well predicted -
in fact, the maximum residuals are observed for the least significant
cliffs.

We then considered the three hold-out molecules and some of the members
of the training set with which they are predicted to show activity
cliffs. For this dataset, there are relatively few ``true'' activity
cliffs. For example, while molecules
\textbf{371259} and \textbf{371307} have a 1000-fold difference in
activity, their $T_c = 0.68$, and would not represent a truly
significant cliff. However, the pair is well poredicted with a
residual of 0.04 log units. if we consider the subset of predictions
where the $T_c \geq 0.8$ we note that the median absolute residual is
0.24 log units. However, the activity differences are not always
significant ranging from 1.1-fold to 13-fold. Two such examples are shown
in Fig.~\ref{fig:dai-comparison}, where the only difference is in the
position of the methyl substituent. Similar behavior is observed with
the other hold-out molecules.

\section{Discussion}
\label{sec:discussion}

The results described above suggest that predicting pairwise SALI
values is a useful way to identify whether a new molecule will
form an activity cliff with one or more members of the training
set. As with all QSAR models, one is limited by the nature of the
underlying data, both in terms of accuracy as well as
applicability. Thus while we have selected a few ChEMBL assays for the
purposes of highlighting this approach, we have found that for some
assays, the model built on pairwise data performs worse than the model
built on the original data. Given that most of the datasets are small
in size (less than 40 molecules), the poor performance of the initial
model is not always surprising. However, in cases where the pairwise
model does not exhibit improved performance, we believe that
predictive modeling of activity cliffs in the original dataset will
not be fruitful.

Given that many of the models show relatively poor statistics on the
pairwise data, it is important to note that much of this is due to the
large variance in the predictions for the less significant cliffs. One
possible reason for the increase in variance for these portions of the
dataset is the fact that the distribution of the log(SALI) values
tends to be skewed to the right (Fig.~\ref{fig:logsalidist}). In
general we see that the smaller SALI values constitute a relatively
small portion of the dataset.

Another factor affecting the models built on pairwise data is the
descriptors that are employed. While the choice of initial descriptors
is always open to debate, we partially avoid a biased selecion by
employing a random forest model to perform implicit feature
selection. However, given an initial set of descriptors we observe
that the models are robust to the different aggregation functions to
derive a pairwise descriptor representation from the descriptors for
the individual compounds. This is a useful observation given that
there appears to be no rule to select one aggregation function over
another \emph{a priori}. While random forests do allow us to avoid
feature selection, it is clear that other models, such as neural
networks or support vector machines, could lead to more predictive
solutions. At this stage, our aim is to highight the modeling
procedure and we believe that the random forest allows us to highight
the utility of this approach without sacrificing too much by way of
predictive performance.

It is important to note that while the modeling approach here focuses
on predicting SALI values rather than actual activities, one can
obtain an idea of the latter based on the (known) activity of the
training set molecules. Thus for a new molecule, predicted to show a
cliff with a member of the training set, we can determine whether this
molecule might be the potent member of the cliff, by noting whether
the corresponding training set molecule is less or more potent. Given
that one is usually interested in activity cliffs because of the
possibility of increasing potency, this observation implies that for a
collection of new molecules, one can prioritize them as being
predicted to be more or less potent. From this point of view, the
approach described here is more suited for prioritization, rather than
prediction of absolute values.

\section{Conclusions}
\label{sec:conclusions}

We have presented an approach to extending a structure-activity
landscape in an indirect fashion, by predicting the propensity of a
new molecule to exhibit an activity cliff with one or more molecules
in a pre-exisiting SAR dataset. The method is based on building
a model on pairwise SALI values (dependent variable) and pairwise
aggregated descriptor values (independent variables). For a new
molecule, we obtain $n$ pairwise SALI predictions (since we must
estimate the SALI value for each training set molecule and the test
molecule). The predicted SALI values can then be used to judge whether
the new molecule will exhibit a cliff and whether such a cliff is in
the desired direction (i.e., improving potency versus worsening
potency). To test this strategy we have developed random forest modls
to predict SALI values for several ChEMBL datasets. While the model
performance statistics are not stellar, we observe that this is
primarily due to high prediction variance at the lower range of SALI
values. In contrast, the more significant cliffs are relatively well
predicted, though, unsurprisingly, the most significant cliffs are not
always well predicted.

In summary, this approach extends the activity cliff concept, from
that of a retrospective analysis tool to a prospective tool that could
be used to guide synthetic campaigns in their goals of improved
potency.

\clearpage
\newpage

\bibliography{paper}

\newpage

\ctable[cap={}, caption={Datasets considered in this study.},
label={tab:data}]
{lp{0.4\textwidth}rrrr}
{}
{\FL
Name & Description & Endpoint & ChEMBL AID & Size & Reference
\ML
Cavalli & \raggedright{Putative hERG inhibitors (via pharmacophore modeling)} & Score &  & 30 & Ref.~\citen{Cavalli:2002aa} \NN
Costanzo & \raggedright{Thrombin inhibitors} & $\mathrm{IC_{50}}$ & 305283 & 60 &
Ref.~\citen{Costanzo:2005uq} \NN
Kalla & \raggedright{A2B adenosine receptor antagonists} & $K_i$ &
364155 & 38 & Ref.~\citen{Kalla:2006kx} \NN
Dai & \raggedright{RTK inhibitors} & $K_i$ & 429056 & 44 & Ref.~\citen{Dai:2007kl}
% Mewshaw & \raggedright{ER$\beta$ inhibitors} & $\mathrm{IC_{50}}$ &
% 306229 & 70 & \cite{Mewshaw:2005vf} \NN
\LL}

\ctable[cap={}, caption={Performance metrics for the random forest
  models built for the datasets used in this study. RMSE and $R^2$ are
based on out-of-bag data\cite{Breiman:1984aa}},
label={tab:rfmodel-results}]
{lrcrr}
{}
{\FL
Dataset & Y-range & Aggregation Function & RMSE & $R^2$ 
\ML
Cavalli & 7.21 & $f_{\textrm{diff}}$ & 1.11& 0.57\NN
        & & $f_{\textrm{mean}}$ & 0.98& 0.67\NN
        & & $f_{\textrm{geom}}$ & 1.01& 0.65\NN \NN
Costanzo & 288.03& $f_{\textrm{diff}}$ & 12.88& 0.24\NN
       & & $f_{\textrm{mean}}$ & 13.56& 0.17\NN \NN
Kalla & 88.25 & $f_{\textrm{diff}}$ & 7.51& 0.62\NN
      &  & $f_{\textrm{mean}}$ & 8.26& 0.53\NN \NN
Dai & 32.71 & $f_{\textrm{diff}}$ & 2.60& 0.44\NN
     &   & $f_{\textrm{mean}}$ & 2.67& 0.41
\LL
}

\clearpage
\newpage

\ctable[cap={}, caption={Distribution of Pearson correlations between
  SALI values and the descriptors for the Cavalli
  dataset. Three descriptor aggregation functions are considered.},
label={fig:desccor1},figure,botcap]
{c}
{}
{
  \includegraphics[width=\linewidth]{sali-desc-corr}
}

\ctable[cap={}, caption={Plots of predicted versus observed SALI values
  obtained using random forest models, on the Cavalli dataset. Each
  panel corresponds to the use of a different descriptor aggregation
  function.}, label={fig:rfmodel-cavalli}, botcap, figure]
{c}
{}
{
\includegraphics[width=\linewidth]{rfmodels-cavalli}
}

\ctable[cap={}, caption={Predicted versus observed SALI values,
  obtained from random forest models for the three ChEMBL datasets. The
 three plots correspond to the two different aggregation functions
 ($f_{\textrm{diff}}$ and $f_{\textrm{mean}}$ respectively).},
label={fig:rfmodel-chembl},botcap,figure]
{p{0.5in}c}
{}
{
\imagetop{\textbf{Costanzo}} & \imagetop{\includegraphics[width=0.9\linewidth]{rfmodels-costanzo}}\NN
\imagetop{\textbf{Kalla}}  & \imagetop{\includegraphics[width=0.9\linewidth]{rfmodels-kalla}} \NN
\imagetop{\textbf{Dai}} &
\imagetop{\includegraphics[width=0.9\linewidth]{rfmodels-dai}}   
}

\ctable[cap={}, caption={Results of random forest models developed
  using the Costanzo dataset. The left plot shows the predicted versus
  observed plot for the model using the original $\mathrm{IC_{50}}$ values, and
  the right hand plot shows the predicted versus observed log(SALI)
  values.}, label={fig:costanzo-pred}, figure, botcap]
{c}
{}
{\includegraphics[width=\linewidth]{pred305283}}

\ctable[cap={}, caption={Results of random forest models developed
  using the Kalla dataset. The left plot shows the predicted versus
  observed plot for the model using the original $K_i$ values, and
  the right hand plot shows the predicted versus observed log(SALI)
  values.}, label={fig:kalla-pred}, figure, botcap]
{c}
{}
{\includegraphics[width=\linewidth]{pred364155}}

\ctable[cap={}, caption={Results of random forest models developed
  using the Dai dataset. The left plot shows the predicted versus
  observed plot for the model using the original $K_i$ values, and
  the right hand plot shows the predicted versus observed log(SALI)
  values.}, label={fig:dai-pred}, figure, botcap]
{c}
{}
{\includegraphics[width=\linewidth]{pred429056}}

\ctable[cap={}, caption={Detailed analysis of SALI predictions for the
  Kalla dataset. \textbf{A} - a plot of predicted versus observed
  log(SALI) values for the training set and the hold out
  set. \textbf{B} - a summary of the training set, where we plot the
  structural difference versus the
  logarithm of the ratio of the activities for each pair of molecules
  in the prediction set. Points are shaded by their absolute
residual. \textbf{C} - a box plot summarizing the distribution of  
residuals associated with predictions from each of the hold out molecules.},
label={fig:tset-pset-kalla}, figure, botcap]
{c}
{}
{\includegraphics[width=\linewidth]{tset-pset-kalla}}

\ctable[cap={}, caption={The hold out molecules for the Kalla dataset
  and training set members with which the hold outs exhibit predicted
  activity cliffs. Bold numbers are ChEMBL MOLREGNO values and numbers
  in parentheses are the absolute prediction
  residual in log(SALI) units.}, figure, botcap,
label={fig:kalla-comparison}]    
{c}
{}
{\includegraphics[width=\linewidth]{kalla-comparison}}

\ctable[cap={}, caption={Detailed analysis of SALI predictions for the
  Dai dataset. \textbf{A} - a plot of predicted versus observed
  log(SALI) values for the training set and the hold out
  set. \textbf{B} - a summary of the training set, where we plot the
  structural difference versus the logarithm of the ratio of the
  activities for each pair of molecules
  in the prediction set. Points are shaded by their absolute
  residual. \textbf{C} - a box plot summarizing the distribution of  
  residuals associated with predictions from each of the hold out
  molecules.}, figure, botcap,
label={fig:tset-pset-dai}]    
{c}
{}
{\includegraphics[width=\linewidth]{tset-pset-dai}}

\ctable[cap={}, caption={The hold out molecules for the Dai dataset
  and training set members with which the hold outs exhibit predicted
  activity cliffs. Bold numbers are ChEMBL MOLREGNO values and numbers
  in parentheses are the absolute prediction
  residual in log(SALI) units.}, figure, botcap,
label={fig:dai-comparison}]    
{c}
{}  
{\includegraphics[width=\linewidth]{dai-comparison}}

\ctable[cap={}, caption={Distribution of log(SALI) values for the
  three ChEMBL datasets.}, figure, botcap,
label={fig:logsalidist}]    
{c}
{}  
{\includegraphics[width=\linewidth]{logsalidist}}

\end{document}