### 2012-07-18 rebuttal.tex
Added more rebuttal

\documentclass[letterpaper, 12pt]{article}
\usepackage{url}
\usepackage{ctable}
\usepackage{amsmath,amssymb}      % for \AA etc
\usepackage[onehalfspacing]{setspace}
\usepackage{graphicx}
\usepackage{times}

\DeclareGraphicsExtensions{.eps,.png}
\onehalfspacing

\setlength{\parindent}{0pt}
\setlength{\parskip}{1em}
\usepackage{anysize}

\begin{document}

I would like to thank the reviewers for their thorough comments on the original submissions. I have
prepared a revised manuscript that address the comments and my responses are detailed below.

\fbox{\textbf{Reviewer: 1}}

\textbf{(1) Please, note that there are numerous typographical errors
throughout the manuscript (which is quite disturbing). 
\vskip 1em
(2) Chosen methods and descriptors should at least be referenced. For
example, no references are provided for BCI and CDK.}

The manuscript has been thoroughly proof read and references included
where appropriate

\textbf{(3) The derivation of RF models is essentially not described
  (except of making reference to default parameters settings in R). It
  remains largely unclear how these models were developed. }

The text has been updated to be more specific on the development of
the random forest model.

(4) The
selection of training and test sets should be clearly specified in the
Methods section. For example, how were the RF models trained and
evaluated for which predictions of pair-wise SALI values are reported
in Figure 2? It is hoped that the reported SALI values were not
predicted for the training data.

(5) No support is provided for the
assumption that the RF models “are relatively robust even for datasets
of small size”. In fact, the author later on states several times that
limited model quality might be due to the use of small compound
sets. Appropriate validation will be essential here.

(6) There is
only limited statistical assessment/validation of the
models. Statements made in the text that R2 values are “reasonable”
and the models perform “relatively well” are difficult to
reconcile. Rather, Table 2 indicates that the R2 values were rather
poor in several instances.

 (7) It appears that there are systematic
prediction errors for pairs. At low value ranges (that are not
relevant for activity cliff assessment), SALI values are consistently
over-predicted. By contrast, at high value ranges (that are relevant),
SALI values are consistently under-predicted, in part
significantly. This is not investigated, although the apparent errors
at high value ranges are the perhaps most critical features of the
reported RF models.

(8) In the section “Extending a landscape”,
predictions are finally reported for only three (!) molecules taken
out of each data set. Carefully put, the predictions are
heterogeneous. On the basis of the results reported for this little
bit of evaluation, one would be hard pressed to make a case for the
ability of the approach to predict activity cliffs with any certainty.

\textbf{(9) For the prediction of activity cliffs, SALI values must be
calculated for compound pairs. This is the central idea of the
approach. Accordingly, fingerprint descriptors and potency values must
be combined in some ways. 
\vskip 1em
(9.1.) It is nowhere stated how activity values are treated for
compounds forming a pair. Are activity differences calculated at the
pair level? If so, how are activity differences compared for the
calculation of pair-based SALI scores - as a difference of
differences? It should be noted predicted potency values in Figures
4-6 are essentially all over the place.
}
The entire premise of the proposed method is that one evaluates SALI
values for a dataset that has observed activities. We state
specifically (Equation 1) how the activities are combined when
molecules are considered pairwise.

It's not clear what the reviewer means by \emph{``\ldots how are
  activity differences compared for the calculation of pair-based SALI
  scores \ldots''}. The SALI value itself represents a pair of
molecules and in this calculation, we consider the absolute activity
difference. We do not consider pairs of pairs

\textbf{(9.2.) To obtain a fingerprint descriptor for a pair, the
  author aggregates fingerprints of individual molecules, finally, by
  simple averaging}

In fact, we do not generate pairwise fingerprint
descriptors. Fingerprints are only used to evaluate the similarity
term in the calculation of the SALI value.

\textbf{ For comparisons of pairs, Tanimoto similarity is then
  calculated for averaged fingerprints, which presents an artificial
  assessment of pair similarity (one can easily come up with a few
  hypothetical compound similarity relationships that make this
  assessment questionable at best). As a consequence, activity cliffs
  are considered for which Tanimoto similarity calculations yield
  values between 0.2 and 0.3!}

I believe that these comments are based on the discussion in Section
3.1 and it appears that this section of the paper was not clear.

Firstly, we have never mentioned the averaging of fingerprint
descriptors. Instead, we generate a descriptor vector for a pair of
molecules by averaging the (topological) descriptor vectors of the
individual members of that pair.

The goal of Section 3.1 is to indicate that even though the Y variable
and the X variables encode structural information (the former
indirectly and the latter directly), there is little correlation
between the individual (aggregated, topological) descriptors and the Y
variable (where sructural information is included via
fingerprints). To quantify this, we evaluate the Pearson correlation
between each (aggregated, topological) descriptor and the Y variable
and show that the maximum value is very low ($R^2 < 0.15$).

This discussion does not involve Tanimoto similarities at all

 Thus, on the basis of these similarity calculations, one could never
 decide which level of similarity might be cliff relevant.

\textbf{ In this
 respect, the descriptions of activity cliff by the author are quite
 telling – he speaks of “predicted” versus “true” activity cliffs, of
 “relatively accurate” activity cliffs, and of “most significant
 activity cliffs that are in fact not very significant activity cliffs
 in an absolute sense”, and so on. }

I agree that the wording of this discussion is rather sloppy. The text
has been updated to be more quantitative.

Yet at the same time, the concept of an activity cliff is somewhat
subjective. In many cases, it is akin to saying \emph{``I'll know it's
  an activity cliff when I see it''}. For example, a pair of molecules
exhibiting a $T_c = 0.9$ and 1000x difference in acivity is obviously
an activity clif. Would a pair with $T_C = 0.85$ and 50x difference in
activity not be an activity cliff? From this point of view, I do not
think it unreasonable to refer somewhat qualitatively to the
``cliffness'' of an activity cliff.

Furthermore, there might be many instances of activity and fingerprint similarity relationships between pairs of compounds that might yield high pair-based SALI scores dominated by either the activity or similarity component, without forming a “true” activity cliff. The conclusion of this reviewer is that it is currently not possible to capture activity cliffs with any degree of certainty on the basis of the pair calculations, as reported.
(10) It also remains largely obscure how an activity landscape should be extended following this approach by compounds for which no activity information is available. Should the activity first be predicted by an RF model and then used for pair-based SALI score calculations? One would hope not.

In summary, there will be quite a bit of work required to get this manuscript in shape for publication, both in terms of its presentation and technical content. However, as long as points (9) and (10) are not thoroughly addressed scientifically, this work should not be published.

To give the author the benefit of the doubt -and credit for the underlying idea- this reviewer calls for “Major Revisions”, although the study might as well be considered premature at this point.

Additional Questions:
Please rate the quality of the science reported in this paper (10 - High Quality/ 1 - Low Quality): 4

Please rate the overall importance of this paper to the field of chemical information or modeling (10 - High Importance / 1 - Low Importance): 6

\fbox{\textbf{Reviewer: 2}}

\textbf{1.  The manuscript is poorly written.  There are many typos, grammatical errors, missing
  references, and incomplete descriptions of methods and results.}

The text has been thoroughly proof read to remove typos and incorrect grammar. References have been
added and methods have been described more clearly. It's not clear which descriptions were
incomplete. However, the methodology section now reads more coherently, and some aspects are now
clearer based on other reviewer comments.

\textbf{2.  Qualitative assessments of results are not acceptable.  The authors should give
  statistical evaluations and corresponding metrics for models and their predictions.}

This was an oversight on our part. The use of qualitative terms has been removed and results and
comparisons are quantitatively characterized. For example,  the random forest models built to
predict SALI values are now compared to Y-scrambled models to ensure that the the results are not
purely due to chance.

\textbf{3.  The tests sets are too small to make any meaningful conclusions.}

\textbf{4.  SALI values needs to be determined for pairs of molecules - is that not the essence of
  the methodology?}

Yes. For a given dataset, with measured activity values, one can
indeed evaluate the SALI values. The proposed methodology takes these
``measured'' SALI values for a training set and then predicts the SALI
values for a new molecule with each member of the training set. As
with traditional QSAR models, a complete validation would require one
to measure the activity of the new molecule and evaluate the actual
SALI values. In absence of experimental activities for the test set,
we resort to a training/test split, which is an accepted approach in
QSAR modeling.

\textbf{5.  There is no discussion/direction of how to use the methodology in new applications.}

Section 4.1 in the original manuscript specifically addresses how one
might use the proposed methodology

\end{document}