### 2012-07-27 letter-resubmit.tex
added new submission letter. fixed some minor typos

\documentclass[11pt]{newlfm}
\usepackage{setspace}
\onehalfspacing
\newlfmP{leftmarginsize=.75in,leftmarginskipright=15pt,topmarginskip=2em,%
stdletter,
noheadline,nofootline,dateright,addrfromright,%
orderfromdateto,sigleft}

\nameto{Prof.~William Jorgensen}
\addrto{%\parbox{2.5in}{
Department of Chemistry\\
Yale University\\
PO Box 208107\\
New Haven, CT 06520-8107}

\namefrom{Rajarshi Guha}
\addrfrom{%\parbox{2.5in}{
NIH Center for Advancing Translational Science \\
9800 Medical Center Drive \\
Rockville, MD 20852
}%}
\phonefrom{814 404 5449}

\begin{document}
    \closeline{Yours Sincerely,}
   \greetto{Dear Prof.~Jorgensen,}
    \begin{newlfm}
      I would like to resubmit the article titled ``\emph{Exploring
        Uncharted Territories - Predicting Activty Cliffs in
        Structure-Activity Landscapes}'' (ci-2012-00047k) for consideration in the
      Journal of Chemical Information and Modeling.

A separate document has been included that addresses each of the
reviewers comments.

%  \vskip 0.5em
    \end{newlfm}
\end{document}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: